---
title: "In Class Exercise 06 - Take-home Exercise 2 kickoff"
format: html
author: "Erica Chan"
date: "27 September , 2025"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

```{r}
pacman::p_load(tidyverse, sf, sfdep, tmap, knitr)
```

Import data

```{r}
BusStop <- st_read(dsn='data/BusStopLocation_Aug2025', 'BusStop')

```

```{r}
mpsz_sf <- st_read("data/MasterPlan2019SubzoneBoundaryNoSeaKML.kml") %>% 
  st_zm(drop = TRUE, what = "ZM") %>% st_transform(crs = 3414)
```

```{r}
odbus=read_csv('data/origin_destination_bus_202508.csv')
```

# Prepare Bus Stop data

check crs. `BusStop` EPSG code is incorrect (9001), so need to use st_transform to convert it to 3414.

```{r}
st_crs(BusStop)
st_crs(mpsz_sf)

```

```{r}
BusStop <- st_transform(BusStop, 3414)
st_crs(BusStop)
```

Quick plot to check for correctness

```{r}
qtm(mpsz_sf) +
qtm(BusStop, add=TRUE)
```

There are some bus stops outside of the Singapore boundaries (busses crossing the border to Johor Bahru). We can use st_join to extract bus stops only in Singapore and combine the mpsz sf dataframe with the bus stop sf dataframe. BusStop **must** be the first argument to preserve the point layer.

```{r}
BusStop_in_SG <- st_join(
  BusStop, mpsz_sf,
  join=st_within,
  left=FALSE #remove rows without matches 
)
```

check output again to ensure points outside of Singapore have been removed

```{r}
qtm(mpsz_sf) +
qtm(BusStop_in_SG, add=TRUE)
```

# Prepare hexagon data

Derive the analytical hexagon using st_make_grid. This function allows us to make grids of different shapes (Eg circle, square, etc).

```{r}
hexagon <- st_make_grid(mpsz_sf,
                        cellsize=700, # distance between opposite edges of the hexagon
                        what='polygon',
                        square=FALSE)%>%
  st_as_sf() #convert to tibbular / sf data frame, otherwise output will be in sfc data form
```

hexagon is a single column data frame, there is no id field.

We want to limit the hexagons to areas where there are bus stops.

```{r}
qtm(hexagon)+
qtm(mpsz_sf, add=TRUE) +
qtm(BusStop_in_SG, add=TRUE)
```

filtering hexagons with bus stops

```{r}
hexagon$n_collisions = lengths(st_intersects( #calculates how many  bus stops intersect with hexagons
  hexagon, BusStop_in_SG
))

hexagon <- filter(hexagon, n_collisions>0) #filters out hexagons which do not contain any bus 
```

```{r}
qtm(mpsz_sf, fill='green')+
  qtm(hexagon, add=TRUE)+
  qtm(BusStop_in_SG, size = 0.2, add=TRUE)
```

assign IDs to each hexagon

```{r}
hexagon <- hexagon %>%
  select(,-n_collisions) #remove n_collisions column to tidy up attribute table

hexagon$HEX_ID <- sprintf(
  "H%04d", seq_len(nrow(hexagon)))%>% 
  as.factor()


kable(head(hexagon)) #view the table 
```

# prepare trip generation data

Cleaning the data

```{r}
trips <- odbus %>%
  select(c(ORIGIN_PT_CODE, DAY_TYPE, TIME_PER_HOUR, TOTAL_TRIPS))%>%
  rename(c(BUS_STOP_N=ORIGIN_PT_CODE, HOUR_OF_DAY=TIME_PER_HOUR, TRIPS=TOTAL_TRIPS))
trips$BUS_STOP_N <- as.factor(trips$BUS_STOP_N)
kable(head(trips))
```

Combine bus stop and hexagon layers

```{r}
bs_hex <- st_intersection(
  BusStop, hexagon) %>% 
  st_drop_geometry() %>% #drops hexagon geometry layer
  select (c(BUS_STOP_N, HEX_ID)) #select attributes of interest
          
kable(head(bs_hex))
```

Add HEX_ID to bus trips data

```{r}
trips <- inner_join(trips, bs_hex) 
kable(head(trips))
```
