---
title: "In-Class Exercise"
format: html
author: "Erica Chan"
date: "September 13, 2025"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

Load in R packages
```{r}
pacman::p_load(sf,tmap, tidyverse, httr,
               ggthemes, plotly, terra, spatstat, rvest)
```

::: callout-note
-   **httr** is an R package designed to provide a wrapper for the curl pakcae, customised to the demands of modern web APIs.
-   **dplyr** is a grammar of data manipulation, providing a consistent set of verbs that help to solve the most common data manipulation challenges.
-   **lubridate** :::
:::

save the data into acra_data

```{r}
folder_path <- "data/ACRA"
file_list <- list.files(path = folder_path,
                        pattern = "^ACRA*.*\\.csv$",
                        full.names= TRUE)
acra_data <- file_list %>%
  map_dfr(read_csv)
```

save the data as rds to avoid having to redo all data preparation steps again later

```{r}
write_rds(acra_data,
          "data/rds/acra_data.rds")
```
tidying ACRA data

```{r}
biz56111 <- acra_data %>%
  select(1:24) %>%
  filter(primary_ssic_code== 56111) %>%
  mutate(date=as.Date(registration_incorporation_date),
         YEAR=year(date),
         MONTH_NUM=month(date),
         MONTH_ABBR=month(date, label=TRUE, abbr=TRUE)) %>%
  mutate(postal_code=str_pad(postal_code, width=6, side="left", pad="0")) %>%
  filter(YEAR>=2024)
```

Geocoding

```{r}
postcodes <- unique (biz56111$postal_code)

url <- "https://onemap.gov.sg/api/common/elastic/search"

found <- data.frame()
not_found <- data.frame(postcode = character())

for(pc in postcodes) {
  query <- list(
    searchVal = pc,
    returnGeom = "Y",
    getAddrDetails ="Y",
    pageNum = "1"
    )
  
res <- GET(url, query=query)
json <- content(res)

if (json$found != 0) {
  df <- as.data.frame(json$results, stringsAsFactors=FALSE)
  df$input_postcode <- pc
  found <- bind_rows(found, df)
  } else{
  not_found <- bind_rows(not_found, data.frame(postcode = pc))
  }
}
```


```{r}
biz_56111=biz56111%>%
  left_join(found,
            by=c('postal_code'='POSTAL'))

biz_56111_sf <- st_as_sf(biz_56111,
                         coords=c("X","Y"),
                         crs=3414)
```


