---
title: "In Class Exercise 7"
format: html
author: "Erica Chan"
date: "27 September , 2025"
date-modified: "last-modified"
execute: 
  eval: false
  echo: true
  warning: false
  freeze: true
---

```{r}
pacman::p_load(olsrr, corrplot, ggpubr, 
               sf, sfdep, GWmodel, tmap, 
               tidyverse, gtsummary,
               performance, RColorBrewer, rvest)
```

import data

```{r}
mpsz_sf <- st_read("data/geospatial/MasterPlan2019SubzoneBoundaryNoSeaKML.kml") %>% 
  st_zm(drop = TRUE, what = "ZM") %>% st_transform(crs = 3414)

extract_kml_field <- function(html_text, field_name) {
  if (is.na(html_text) || html_text == "") return(NA_character_)
  
  page <- read_html(html_text)
  rows <- page %>% html_elements("tr")
  
  value <- rows %>%
    keep(~ html_text2(html_element(.x, "th")) == field_name) %>%
    html_element("td") %>%
    html_text2()
  
  if (length(value) == 0) NA_character_ else value
}

mpsz <- mpsz_sf %>%
  mutate(
    REGION_N = map_chr(Description, extract_kml_field, "REGION_N"),
    PLN_AREA_N = map_chr(Description, extract_kml_field, "PLN_AREA_N"),
    SUBZONE_N = map_chr(Description, extract_kml_field, "SUBZONE_N")
    
  ) %>%
  select(-Name, -Description) %>%
  relocate(geometry, .after = last_col())
```

```{r}
tm <- mpsz %>%
  filter(PLN_AREA_N=="TAMPINES")
```

```{r}
condo_resale = read_csv("data/aspatial/Condo_resale_2015.csv")%>%
  st_as_sf(coords = c("LONGITUDE", "LATITUDE"),
                            crs=4326) %>%
  st_transform(crs=3414)
```

## Create GWR with fixed bandwidth and AIC

```{r}
bw.fixed <- bw.gwr(
  formula = SELLING_PRICE ~ AREA_SQM + AGE  + 
    PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE    +
    PROX_URA_GROWTH_AREA + PROX_MRT + PROX_PARK + 
    PROX_PRIMARY_SCH + PROX_SHOPPING_MALL   + 
    PROX_BUS_STOP   + NO_Of_UNITS + FAMILY_FRIENDLY + 
    FREEHOLD, 
  data=condo_resale, 
  approach="AIC", 
  kernel="gaussian", 
  adaptive=FALSE, 
  longlat=FALSE)
```

Compare results with Hands-on Ex07 where we used CV instead of AIC:

-   AIC leads to using a larger fixed bandwidth (1391.309m instead of 971.2405m)

Next, calibrate the model using fixed bandwidth 1391.309

```{r}
gwr.fixed <- gwr.basic(
  formula = SELLING_PRICE ~ AREA_SQM + AGE  + 
    PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE    +
    PROX_URA_GROWTH_AREA + PROX_MRT + PROX_PARK +
    PROX_PRIMARY_SCH + PROX_SHOPPING_MALL   + 
    PROX_BUS_STOP   + NO_Of_UNITS + FAMILY_FRIENDLY + 
    FREEHOLD, 
  data=condo_resale, 
  bw=bw.fixed, 
  kernel = 'gaussian', 
  longlat = FALSE)

gwr.fixed
```

Compare results with Hands-on Ex07 where we used CV instead of AIC:

-   Adjusted R-square: Using AIC gives a lower R-square value (0.8182339 vs 0.890912), which means the model using AIC can explain less variation in the data compared to the model made using CV.

Takeaway: When calibrating GWR, should use both criterions to compare and use the one which gives better results in the model.

## Creating GWR with adaptive bandwidth using AIC

```{r}
bw.adaptive <- bw.gwr(
  formula = SELLING_PRICE ~ AREA_SQM + AGE  + 
    PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE    +
    PROX_URA_GROWTH_AREA + PROX_MRT + PROX_PARK +
    PROX_PRIMARY_SCH + PROX_SHOPPING_MALL   + 
    PROX_BUS_STOP   + NO_Of_UNITS + 
    FAMILY_FRIENDLY + FREEHOLD, 
  data=condo_resale, 
  approach="AIC", 
  kernel="gaussian", 
  adaptive=TRUE, 
  longlat=FALSE)
```

Compare results with Hands-on Ex07 where we used CV instead of AIC:

-   Using AIC recommends fewer number of nearest neighbours than CV (27 vs 30).

```{r}
gwr.adaptive <- gwr.basic(
  formula = SELLING_PRICE ~ AREA_SQM + AGE + 
    PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE +
    PROX_URA_GROWTH_AREA + PROX_MRT + PROX_PARK +
    PROX_PRIMARY_SCH + PROX_SHOPPING_MALL + PROX_BUS_STOP +
    NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD,
  data=condo_resale, 
  bw=bw.adaptive, 
  kernel = 'gaussian', 
  adaptive=TRUE, 
  longlat = FALSE)

gwr.adaptive
```

Compare results with Hands-on Ex07 where we used CV instead of AIC:

-   Adjusted R-square: Using AIC gives a slightly higher R-square value (0.8634199 vs 0.8561185), which means the model using AIC can explain slightly more variation in the data compared to the model made using CV.

Final conclusion: Across all 4 models, the model with the highest explanatory value is the gwr model using fixed bandwidth and CV (R-square=0.890912). Hence, this is model should be chosen.

## Visualising local R-square values in Tampines Planning Area

```{r}
bw.fixed <- bw.gwr(
  formula = SELLING_PRICE ~ AREA_SQM + AGE  + 
    PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE    +
    PROX_URA_GROWTH_AREA + PROX_MRT + PROX_PARK + 
    PROX_PRIMARY_SCH + PROX_SHOPPING_MALL   + 
    PROX_BUS_STOP   + NO_Of_UNITS + FAMILY_FRIENDLY + 
    FREEHOLD, 
  data=condo_resale, 
  approach="CV", 
  kernel="gaussian", 
  adaptive=FALSE, 
  longlat=FALSE)

gwr.fixed <- gwr.basic(
  formula = SELLING_PRICE ~ AREA_SQM + AGE  + 
    PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE    +
    PROX_URA_GROWTH_AREA + PROX_MRT + PROX_PARK +
    PROX_PRIMARY_SCH + PROX_SHOPPING_MALL   + 
    PROX_BUS_STOP   + NO_Of_UNITS + FAMILY_FRIENDLY + 
    FREEHOLD, 
  data=condo_resale, 
  bw=bw.fixed, 
  kernel = 'gaussian', 
  longlat = FALSE)

condo_resale.sf.fixedcv <-
  st_as_sf(gwr.fixed$SDF) %>%
  st_transform(crs=3414)
```

filter for points within planning area tampines

```{r}
condo_resale.tm <- st_intersection(condo_resale.sf.fixedcv, tm)

glimpse(conde_resale.tm)
```

plot the map

```{r}
tmap_mode("plot")

tm_shape(tm)+
  tm_polygons(fill_alpha = 0.1) +
tm_shape(condo_resale.tm) +  
  tm_dots(col = "Local_R2",
          border.col = "gray60",
          border.lwd = 1) 
```

There are 76 data points in tampines representing resales of condos in this planning area. The local R-square of these data points is relatively high ranging from 0.84 to 0.98, indicating that the gwr model is able to explain most of the variation in these data points. However, we can see one cluster of datapoints where the R-square value is relatively lower (in the Tampines West subzone). This may indicate that the model is relatively poorer at explaining the variation of prices of the condominiums in this cluster/subzone, and that there are other unidentified factors influencing price of resold condominiums here. On the other hand, all other data points in Tampines have relatively high R-square value, indicating the model can explian most of the price variation in these data points.
