---
title: "In-class Exercise 08"
format: html
author: "Erica Chan"
date: "25 October , 2025"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

```{r}
pacman::p_load(httr, tidyverse, sf, tmap, jsonlite, progress,
               spdep, GWmodel, SpatialML, rsample, Metrics,
               knitr, spatialRF, randomForestExplainer)
```

-   httr - allows HTTP calls to OneMap API
-   jsonlite - read json files to tibbular data frame
-   sf - handles geospatial data
-   tmap - handles maps
-   progress - tracks computations

## geocoding HDB resale data

```{r}
HDBresale <- read_csv("data/rawdata/HDBresale.csv") %>%
                         filter(flat_type=="4 ROOM") %>%
                         filter(month >= "2025-08" & month <="2025-10")
```

```{r}
HDBresale$street_name<-gsub(
  "ST\\.", "SAINT", 
  HDBresale$street_name
)
```

```{r}
geocode <- function(block, streetname) {
  base_url="https://onemap.gov.sg/api/common/elastic/search"
  address<- paste(block, streetname, sep=" ")
  query <- list("searchVal" = address, 
                "returnGeom" = "Y",
                "getAddrDetails" = "N",
                "pageNum" = "1")
  
  res <- GET(base_url,  query = query)
  restext <- content(res, as="text")
  
  output <- fromJSON(restext) %>%
    as.data.frame() %>%
    select(results.LATITUDE, results.LONGITUDE)
  
  return(output)
}
```

```{r}
#| eval: false
HDBresale$LATITUDE <- 0
HDBresale$LONGITUDE <- 0

for (i in 1:nrow(HDBresale)){
  temp_output <- geocode(HDBresale[i,4], HDBresale[i,5])
  HDBresale$LATITUDE[i] <- temp_output$results.LATITUDE
  HDBresale$LONGITUDE[i] <- temp_output$results.LONGITUDE
}
```

```{r}
#| eval: false
write_rds(HDBresale, "data/rds/HDBresale.rds")
```

```{r}
HDBresale<-read_rds('data/rds/HDBresale.rds')

HDBresale_sf <- st_as_sf(HDBresale, 
                         coords = c("LONGITUDE", "LATITUDE"), 
                         crs = 4326) %>%
  st_transform(crs=3414)
```

```{r}
preschool=st_read("data/rawdata/PreSchoolsLocation.kml")%>%
  st_transform(crs=3414)
```

```{r}
within_350m <- st_is_within_distance(
  HDBresale_sf, preschool, dist=350) #measurement is in meters

HDBresale_sf <- HDBresale_sf %>% 
  mutate(preschool_within350m = map_int(within_350m, length))
```

# modelling

```{r}
mdata <- read_rds("data/rawdata/mdata.rds")
```

data sampling - prorotype models with 10% sample of data. Selecting sample will be done by random using code chunk below

```{r}
set.seed(1234)
HDB_sample <- mdata %>%
  sample_n(1500)
```

## checking of overlapping points

when using GWmodel to calibrate explanatory or predictive models, it is important to ensure no overlapping point features

```{r}
overlapping_points<- HDB_sample %>%
  mutate(overlap=lengths(st_equals(., .))>1)
```

## spatial jittter

jthe code chunk below uses st_jitter of **sf** package to move point features by 1m to avoid overlapping point features

```{r}
HDB_sample<-HDB_sample %>%
  st_jitter(amount=1)
```

## Data sampling

`initial_sample` of **rsample** package allows splitting of data into test and training sets.

```{r}
set.seed(1234)
resale_split <- initial_split(HDB_sample, prop = 6.67/10)
train_data <- training(resale_split)
test_data <- testing(resale_split)
```

## model calibration

```{r}
#| eval: false
bw_adaptive <- bw.gwr(
  formula = resale_price ~ floor_area_sqm +
    storey_order + remaining_lease_mths +
    PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
    PROX_MRT + PROX_PARK + PROX_MALL + 
    PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
    WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
    WITHIN_1KM_PRISCH,
  data=train_data,
  approach="CV",
  kernel="gaussian",
  adaptive=TRUE,
  longlat=FALSE)
```

```{r}
#| eval: false
write_rds(bw_adaptive,"data/model/bw_adaptive.rds")
```

```{r}
#| echo: false
bw_adaptive <- read_rds("data/model/bw_adaptive.rds")
```

```{r}
#| eval: false
gwr_adaptive <- gwr.basic(
  formula = resale_price ~ floor_area_sqm +
    storey_order + remaining_lease_mths + 
    PROX_CBD + PROX_ELDERLYCARE + 
    PROX_HAWKER + PROX_MRT + PROX_PARK + 
    PROX_MALL + PROX_SUPERMARKET +
    WITHIN_350M_KINDERGARTEN +
    WITHIN_350M_CHILDCARE + 
    WITHIN_350M_BUS + WITHIN_1KM_PRISCH,
  data=train_data,
  bw=bw_adaptive, 
  kernel = 'gaussian', 
  adaptive=TRUE,
  longlat = FALSE)

write_rds(gwr_adaptive, 
          "data/model/gwr_adaptive.rds")
```

```{r}
#| echo: false

gwr_adaptive <- read_rds("data/model/gwr_adaptive.rds")
```

```{r}
gwr_adaptive
```

First, p-value should be small to ensure the model is significant.

compare global and local model with adjusted r-square value, then AICc

base (global) model should have adjusted r-square value of 0.6 and above to ensure it explains variation in dependent variable

for local model, there is a significant improvement in adjusted R-square value. Justifies incorporation of explicit location and geographical data. Properties closer together are likely to be more similar than properties further apart.
