---
title: "In-Class_Ex10"
format: html
---

# Import R Packages

-   sf for handling geospatial data and performing geoprocessing function such as computing distance matrix,
-   MCI for performing Huff and MCI modelling,
-   tidyverse for performing data science tasks including visualisation,
-   ggstatsplot, an R package that extends the ggplot2 visualization package to combine data visualization with statistical analysis, creating plots that include statistical test results directly, and
-   tmap for plotting cartographic quality thematic maps.

```{r}
pacman::p_load(sf, tidyverse, MCI, tmap, ggstatsplot)
```

# Import data

## Geospatial data

```{r}
stores <- st_read(
  dsn = "data/geospatial",
  layer = "STORES_TC")

town <- st_read(
  dsn = "data/geospatial",
  layer = "TOWN_TC")

village <- st_read(
  dsn = "data/geospatial",
  layer = "VILLAGE_TC")
```

```{r}
tmap_mode("view")
tm_shape(town) +
  tm_polygons() +
tm_shape(village) +
  tm_polygons() + 
tm_shape(stores) +
  tm_dots()
```

The output print above shown that the three data sets are in **TWD97 / TM2 zone 121 (i.e epsg: 3826)** projected coordinates systems.

```{r}
tmap_mode('plot')
```

## Aspatial data

```{r}
pop2019 <- read_csv(
  "data/aspatial/pop2019.csv")

store_sales <- read_csv(
  "data/aspatial/Store_sales_data.csv")

summary(store_sales)

```

::: callout-note
`store Code` in `store_sales` is matched to `store_CD` in `stores` (different column name for matched values)

`V-ID` in `pop2019` is matched to `VILLCODE` in `village` , but also needs additional cleaning to remove '-' character
:::

# Data preparation

```{r}
store_sales_aggre <- store_sales %>%
  group_by(`Store Code`) %>%
  summarise(
    "Total_Bill" = sum(Bills, na.rm = TRUE), 
    "Total_Sales" = sum(Amount, na.rm = TRUE))
#na.rm=TRUE ensures null values are accepted and will not cause errors
```

```{r}
stores <- stores %>%
  left_join(store_sales_aggre,
            by = join_by(
              Store_CD == `Store Code`)) %>%
  mutate(Store_size = 100) 
```

`summary()` is then used to check for any missing values in the data frame.

```{r}
summary(stores)
```

## Tidy pop2019 data

```{r}
pop2019 <- pop2019 %>%
  mutate(V_ID = str_remove_all(
    V_ID, "-"))
```

data in pop2019 data.frame will be appended on the village sf data.frame by using `left_join()` of **dplyr** package

```{R}
village <- village %>%
  left_join(pop2019,
            by = join_by(
              VILLCODE == V_ID))
```

```{r}
summary(village)
```

COUNTY_ID, TOWN_ID and the counts (eg H_COUNT, P_COUNT...) contain 5 missing values.

This will cause issues for the Huff model computation later, hence we should filter out the NA values.

```{r}
village <- village %>%
  filter(!is.na(P_CNT)) 
```

# Compute distance matrix

```{r}
village_center <- st_point_on_surface(village)
distmat <- st_distance(village_center, stores) 
```

::: callout-note
Using `st_point_on_surface` instead of `st_centroid` guarantees that the returned point lies **inside or on the boundary** of the input geometry
:::

```{r}
rownames(distmat) <- village$`VILLCODE`
colnames(distmat) <- stores$`Store_CD`

distmat_long <- distmat %>%
  as.data.frame() %>%
  rownames_to_column(var = "VILLCODE") %>%
  pivot_longer(
    cols = -VILLCODE,
    names_to = "Store_CD",
    values_to = "distance"
  ) %>%
  mutate(distance = as.numeric(distance)) #as.numeric ensures the 'm' from distmat is ignored
```

# Preparing interaction matrix

```{r}
interaction_matrix <- distmat_long %>%
  left_join(
    stores %>% 
      st_drop_geometry() %>% 
      select(Store_CD, Store_size),
    by = "Store_CD") %>%
  relocate(
    distance, .after = last_col()) 
```

```{r}
summary(interaction_matrix)
```

# Computing Huff Model

```{r}
huff_share <- huff.shares(
  interaction_matrix, "VILLCODE", "Store_CD", 
  "Store_size", "distance")
# default: alpha=1, beta=2
```

## Trade area analysis for a single store

select market share data of a store

```{r}
market_share <- huff_share %>%
  filter(Store_CD == "TG")
```

derive trade areas of the selected store

```{r}
Trade_area <- village %>%
  left_join(market_share)
```

Extract the selected store from stores sf data.frame.

```{r}
selected_store <- stores %>%
  filter(Store_CD == "TG")
```

```{r}
#| fig-width: 10
tm_shape(Trade_area) +
  tm_polygons(fill = "p_ij",
              fill.scale = tm_scale_intervals(
                style = "quantile", 
                n = 10,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                title = "Patronage Probability")) +
  tm_shape(selected_store) +
    tm_dots(size = 0.3,
            fill = "red",
            col = "black") +
  tm_title("Distribution of Patronage Probability of Store TG by village") +
  tm_layout(frame = TRUE) +
  tm_borders(fill_alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scalebar() +
  tm_grid(alpha =0.2) 
```

## Store Sales Analysis

```{r}
huff_share_pp <- huff_share %>%
  left_join(village %>% 
      st_drop_geometry() %>% 
      select(VILLCODE, P_CNT),
    by = "VILLCODE") %>%
  relocate(
    distance, .after = last_col())
```

calculate total sales

```{r}
huff_total <- shares.total(
  huff_share_pp, "VILLCODE", "Store_CD",
  "p_ij", "P_CNT")
```

append estimated sales to stores dataframe

```{r}
store_sales_est <- stores %>%
  left_join(huff_total,
            by = join_by(
              "Store_CD" == "suppliers_single"))
```

```{r}
store_sales_nongeom <- store_sales_est %>% 
  st_drop_geometry()

ggscatterstats(
  store_sales_nongeom,
  x = sum_E_j,
  y = Total_Sales)
```

```{r}
huff_share_wider <- huff_share %>%
  select(VILLCODE, Store_CD, p_ij) %>%
    pivot_wider(names_from = Store_CD, 
                values_from = p_ij)
```
