---
title: "Take-Home_Ex03"
format: html
---

```{r}
pacman::p_load(httr, tidyverse, sf, tmap, jsonlite, progress,
               spdep, GWmodel, SpatialML, rsample, Metrics,
               knitr, spatialRF, randomForestExplainer,
               rvest, stringr, lubridate)
```

-   httr - allows HTTP calls to OneMap API
-   jsonlite - read json files to tibbular data frame
-   sf - handles geospatial data
-   tmap - handles maps
-   progress - tracks computations

## geocoding HDB resale data

```{r}
HDBresale <- read_csv("data/hdb/HDBresale.csv") %>%
                         filter(flat_type=="4 ROOM") %>%
                         filter(month >= "2024-07" & month <="2025-09")
```

```{r}
HDBresale$street_name<-gsub(
  "ST\\.", "SAINT", 
  HDBresale$street_name
)
```

```{r}
geocode <- function(block, streetname) {
  base_url="https://onemap.gov.sg/api/common/elastic/search"
  address<- paste(block, streetname, sep=" ")
  query <- list("searchVal" = address, 
                "returnGeom" = "Y",
                "getAddrDetails" = "N",
                "pageNum" = "1")
  
  res <- GET(base_url,  query = query)
  restext <- content(res, as="text")
  
  output <- fromJSON(restext) %>%
    as.data.frame() %>%
    select(results.LATITUDE, results.LONGITUDE)
  
  return(output)
}
```

```{r}
#| eval: false
HDBresale$LATITUDE <- 0
HDBresale$LONGITUDE <- 0

for (i in 1:nrow(HDBresale)){
  temp_output <- geocode(HDBresale[i,4], HDBresale[i,5])
  HDBresale$LATITUDE[i] <- temp_output$results.LATITUDE
  HDBresale$LONGITUDE[i] <- temp_output$results.LONGITUDE
}
```

```{r}
#| eval: false
write_rds(HDBresale, "data/rds/HDBresale.rds")
```

```{r}
HDBresale<-read_rds('data/rds/HDBresale.rds')

HDBresale_sf <- st_as_sf(HDBresale, 
                         coords = c("LONGITUDE", "LATITUDE"), 
                         crs = 4326) %>%
  st_transform(crs=3414)
```

## Preparing attribute data

::: panel-tabset
### Storey level

```{r}
storey_mapping <- c("01 TO 03" = 1, "04 TO 06" = 2, "07 TO 09" = 3, 
                    "10 TO 12" = 4, "13 TO 15" = 5, "16 TO 18" = 6,
                    "19 TO 21" = 7, "22 TO 24" = 8, "25 TO 27" = 9,
                    "28 TO 30" = 10, "31 TO 33" = 11, "34 TO 36" = 12,
                    "37 TO 39" = 13, "40 TO 42" = 14, "43 TO 45" = 15, 
                    "46 TO 48" = 16, "49 TO 51" = 17)
HDBresale_sf$storey_order <- storey_mapping[HDBresale_sf$storey_range]
```

### Remaining Lease

```{r}
HDBresale_sf <- HDBresale_sf %>%
  mutate(
    years = as.numeric(str_extract(remaining_lease, "^\\d+")),
    months = as.numeric(str_extract(remaining_lease, "\\d+(?=\\s+month)")) %>%
      replace_na(0),
    remaining_lease_months = (years * 12) + months
  )
```

### Unit age

```{r}
HDBresale_sf <- HDBresale_sf %>%
  mutate(
    unit_age_months = (99 * 12) - remaining_lease_months
  )
```

### Bus Stop

```{r}
busstop <- fromJSON("data/transport/BusStops.json") %>%
  as.data.frame

busstop_sf <- st_as_sf(busstop, 
                         coords = c("value.Longitude", "value.Latitude"), 
                         crs = 4326) %>%
  st_transform(crs=3414)%>%
  select(value.BusStopCode, geometry)
```

```{r}
bs_within_350m <- st_is_within_distance(
  HDBresale_sf, busstop_sf, dist=350) #measurement is in meters

HDBresale_sf <- HDBresale_sf %>% 
  mutate(bus_within350m = map_int(bs_within_350m, length))
```

### MRT

```{r}
mrt <- st_read('data/transport/LTAMRTStationExitKML.kml') %>%
  st_transform(crs=3414)

HDBresale_sf <- HDBresale_sf %>%
  mutate(
    prox_mrt = as.numeric(
      st_distance(geometry,
                  mrt[st_nearest_feature(
                    geometry, mrt), ], 
                  by_element = TRUE)
      )/1000
    )
```

```{r}
mrt_within_350m <- st_is_within_distance(
  HDBresale_sf, mrt, dist=350) #measurement is in meters

HDBresale_sf <- HDBresale_sf %>% 
  mutate(mrt_within350m = map_int(mrt_within_350m, length))
```

### Cycling Path

```{r}
cycling <- st_read('data/transport/CyclingPathNetworkKML.kml') %>%
  st_transform(crs=3414)

HDBresale_sf <- HDBresale_sf %>%
  mutate(
    prox_cycling = as.numeric(
      st_distance(geometry,
                  cycling[st_nearest_feature(
                    geometry, cycling), ], 
                  by_element = TRUE)
      )/1000
    )
```

### Taxi Stop

```{r}
taxi <- st_read('data/transport/LTATaxiStopKML.kml') %>%
  st_transform(crs=3414)

HDBresale_sf <- HDBresale_sf %>%
  mutate(
    prox_taxi = as.numeric(
      st_distance(geometry,
                  taxi[st_nearest_feature(
                    geometry, taxi), ], 
                  by_element = TRUE)
      )/1000
    )

t_within_350m <- st_is_within_distance(
  HDBresale_sf, taxi, dist=350) #measurement is in meters
```

### Pre-schools

```{r}
preschool <- st_read("data/family-facility/PreSchoolsLocation.kml")%>%
  st_transform(crs=3414)

within_350m <- st_is_within_distance(
  HDBresale_sf, preschool, dist=350) #measurement is in meters

HDBresale_sf <- HDBresale_sf %>% 
  mutate(preschool_within350m = map_int(within_350m, length))
```

### Kindergartens

```{r}
kindergartens <- st_read("data/family-facility/Kindergartens.kml") %>%
  st_transform(crs=3414)
```

```{r}
k_within_350m <- st_is_within_distance(
  HDBresale_sf, kindergartens, dist=350) #measurement is in meters

HDBresale_sf <- HDBresale_sf %>% 
  mutate(kindergarten_within350m = map_int(k_within_350m, length))
```

### Child Care Services

```{r}
childcare <- st_read("data/family-facility/ChildCareServices.kml") %>%
  st_transform(crs=3414)

c_within_350m <- st_is_within_distance(
  HDBresale_sf, childcare, dist=350) #measurement is in meters

HDBresale_sf <- HDBresale_sf %>% 
  mutate(childcare_within350m = map_int(c_within_350m, length))
```

### Primary Schools

```{r}
#| eval: false
prischool=read_csv('data/family-facility/Generalinformationofschools.csv')
```

```{r}
#| eval: false
prischool$LATITUDE <- 0
prischool$LONGITUDE <- 0

for (i in 1:nrow(prischool)){
  temp_output <- geocode(prischool[i,3],'')
  prischool$LATITUDE[i] <- temp_output$results.LATITUDE
  prischool$LONGITUDE[i] <- temp_output$results.LONGITUDE
}
```

```{r}
#| eval: false
write_rds(prischool, 'data/rds/prischool.rds')
```

```{r}
prischool<-read_rds('data/rds/prischool.rds')

prischool_sf <- st_as_sf(prischool, 
                         coords = c("LONGITUDE", "LATITUDE"), 
                         crs = 4326) %>%
  st_transform(crs=3414)
```

```{r}
p_within_1km <- st_is_within_distance(
  HDBresale_sf, prischool_sf, dist=1000) #measurement is in meters

HDBresale_sf <- HDBresale_sf %>% 
  mutate(prischool_within1km = map_int(p_within_1km, length))
```

### Student Care Services

```{r}
studentcare <- st_read("data/family-facility/StudentCareServicesKML.kml") %>%
  st_transform(crs=3414)

sc_within_350m <- st_is_within_distance(
  HDBresale_sf, studentcare, dist=350) #measurement is in meters

HDBresale_sf <- HDBresale_sf %>% 
  mutate(studentcare_within350m = map_int(sc_within_350m, length))
```

```{r}
HDBresale_sf <- HDBresale_sf %>%
  mutate(
    prox_studentcare = as.numeric(
      st_distance(geometry,
                  studentcare[st_nearest_feature(
                    geometry, studentcare), ], 
                  by_element = TRUE)/1000 #divide by 1000 to get measurement in km
      )
    )
```

### Elder Care Services

```{r}
eldercare <- st_read('data/family-facility/EldercareServicesKML.kml') %>%
  st_transform(crs=3414)

HDBresale_sf <- HDBresale_sf %>%
  mutate(
    prox_eldercare = as.numeric(
      st_distance(geometry,
                  eldercare[st_nearest_feature(
                    geometry, eldercare), ], 
                  by_element = TRUE)
      )/1000
    )
```

### CBD

```{r}
mpsz <- read_rds('data/rds/mpsz_cl.rds')

cbd <- mpsz %>%
  filter(PLN_AREA_N=="DOWNTOWN CORE")%>%
  st_union()

HDBresale_sf <- HDBresale_sf %>%
  mutate(
    prox_cbd = as.numeric(
      st_distance(geometry,
                  cbd[st_nearest_feature(
                    geometry, cbd), ], 
                  by_element = TRUE)
      )/1000
    )
```

### CHAS clinics

```{r}
chas <- st_read('data/facilities/CHASClinics.kml') %>%
  st_transform(crs=3414)

HDBresale_sf <- HDBresale_sf %>%
  mutate(
    prox_CHAS = as.numeric(
      st_distance(geometry,
                  chas[st_nearest_feature(
                    geometry, chas), ], 
                  by_element = TRUE)
      )/1000
    )
```

### Eating Establishments

```{r}
food <- st_read("data/facilities/EatingEstablishments.kml") %>%
  st_transform(crs=3414)

f_within_350m <- st_is_within_distance(
  HDBresale_sf, food, dist=350) #measurement is in meters

HDBresale_sf <- HDBresale_sf %>% 
  mutate(foodestablishments_within350m = map_int(f_within_350m, length))
```

### Exercise Park

```{r}
ep <- st_read('data/facilities/ParksSGKML.kml') %>%
  st_transform(crs=3414)

HDBresale_sf <- HDBresale_sf %>%
  mutate(
    prox_exercisepark = as.numeric(
      st_distance(geometry,
                  ep[st_nearest_feature(
                    geometry, ep), ], 
                  by_element = TRUE)
      )/1000
    )
```

### Gyms

```{r}
gyms <- st_read('data/facilities/GymsSGKML.kml') %>%
  st_transform(crs=3414)

HDBresale_sf <- HDBresale_sf %>%
  mutate(
    prox_gym = as.numeric(
      st_distance(geometry,
                  gyms[st_nearest_feature(
                    geometry, gyms), ], 
                  by_element = TRUE)
      )/1000
    )
```

### Hawker Centres

```{r}
hawker <- st_read('data/facilities/HawkerCentresKML.kml') %>%
  st_transform(crs=3414)

HDBresale_sf <- HDBresale_sf %>%
  mutate(
    prox_hawker = as.numeric(
      st_distance(geometry,
                  hawker[st_nearest_feature(
                    geometry, hawker), ], 
                  by_element = TRUE)
      )/1000
    )
```

### Hospitals

```{r}
hospitals <- read_csv('data/facilities/hospitals_with_coordinates.csv')
hospitals_sf <- st_as_sf(hospitals, 
                         coords = c("longitude", "latitude"), 
                         crs = 4326) %>%
  st_transform(crs=3414)
```

```{r}
HDBresale_sf <- HDBresale_sf %>%
  mutate(
    prox_hospital = as.numeric(
      st_distance(geometry,
                  hospitals_sf[st_nearest_feature(
                    geometry, hospitals_sf), ], 
                  by_element = TRUE)
      )/1000
    )
```

### Libraries

```{r}
library <- st_read('data/facilities/Libraries.kml') %>%
  st_transform(crs=3414)

HDBresale_sf <- HDBresale_sf %>%
  mutate(
    prox_library = as.numeric(
      st_distance(geometry,
                  library[st_nearest_feature(
                    geometry, library), ], 
                  by_element = TRUE)
      )/1000
    )
```

### NEA Market/Food Centres

```{r}
market <- st_read('data/facilities/NEAMarketandFoodCentreKML.kml') %>%
  st_transform(crs=3414)

HDBresale_sf <- HDBresale_sf %>%
  mutate(
    prox_market = as.numeric(
      st_distance(geometry,
                  market[st_nearest_feature(
                    geometry, market), ], 
                  by_element = TRUE)
      )/1000
    )
```

### Parks

```{r}
parks <- st_read('data/facilities/Parks.geojson') %>%
  st_transform(crs=3414)

HDBresale_sf <- HDBresale_sf %>%
  mutate(
    prox_parks = as.numeric(
      st_distance(geometry,
                  parks[st_nearest_feature(
                    geometry, parks), ], 
                  by_element = TRUE)
      )/1000
    )
```

### Pharmacies

```{r}
pharmacy <- st_read('data/facilities/RetailpharmacylocationsKML.KML') %>%
  st_transform(crs=3414)

HDBresale_sf <- HDBresale_sf %>%
  mutate(
    prox_pharmacy = as.numeric(
      st_distance(geometry,
                  pharmacy[st_nearest_feature(
                    geometry, pharmacy), ], 
                  by_element = TRUE)
      )/1000
    )
```

### Shopping Malls

```{r}
malls <- read_csv('data/facilities/singapore_malls_pois.csv') 

malls_sf <- st_as_sf(malls, 
                         coords = c("lon", "lat"), 
                         crs = 4326) %>%
  st_transform(crs=3414)

HDBresale_sf <- HDBresale_sf %>%
  mutate(
    prox_mall = as.numeric(
      st_distance(geometry,
                  malls_sf[st_nearest_feature(
                    geometry, malls_sf), ], 
                  by_element = TRUE)
      )/1000
    )
```

### Supermarkets

```{r}
supermarket <- st_read('data/facilities/SupermarketsKML.KML') %>%
  st_transform(crs=3414)

HDBresale_sf <- HDBresale_sf %>%
  mutate(
    prox_supermarket = as.numeric(
      st_distance(geometry,
                  supermarket[st_nearest_feature(
                    geometry, supermarket), ], 
                  by_element = TRUE)
      )/1000
    )
```

### Neighbourhood Renewal Programme

```{r}
nrp <- st_read('data/hdb/NRPKML.kml') %>%
  st_transform(crs=3414)
```

```{r}
extract_kml_field <- function(html_text, field_name) {
  if (is.na(html_text) || html_text == "") return(NA_character_)
  
  page <- read_html(html_text)
  rows <- page %>% html_elements("tr")
  
  value <- rows %>%
    keep(~ html_text2(html_element(.x, "th")) == field_name) %>%
    html_element("td") %>%
    html_text2()
  
  if (length(value) == 0) NA_character_ else value
}
```

```{r}
nrp_sf <- nrp %>%
  mutate(
    address = map_chr(Description, extract_kml_field, "NAME"),
    est_cmpltn = map_chr(Description, extract_kml_field, "ESTMT_CNSTRN_CMPLTN")
  ) %>%
  select(-Name, -Description) %>%
  mutate(
    quarter_num = as.numeric(str_extract(est_cmpltn, "^\\d")),
    year_num = str_extract(est_cmpltn, "\\d{4}$"),
    end_month = case_when(
      quarter_num == 1 ~ "03", # Q1 ends in March
      quarter_num == 2 ~ "06", # Q2 ends in June
      quarter_num == 3 ~ "09", # Q3 ends in September
      quarter_num == 4 ~ "12", # Q4 ends in December
      TRUE ~ NA_character_
    ),
    cmpltn_date = paste0(year_num, "-", end_month)
  ) %>%
  # Select the original column and the final date object
  select(address, cmpltn_date, geometry)
```

```{r}
HDBresale_sf <- HDBresale_sf %>%
  st_join(nrp_sf %>% select(cmpltn_date), left = TRUE) %>%
  mutate(
    nrp_completed = case_when(
      !is.na(cmpltn_date) & (month > cmpltn_date) ~ 1,
      TRUE ~ 0
    )
  ) %>%
  select(-cmpltn_date)
```
:::

### Cleaning up table

```{r}
HDBresale_sf <- HDBresale_sf %>%
  select(month, resale_price, floor_area_sqm, storey_order, 
         remaining_lease_months, prox_mrt, prox_cycling,
         prox_taxi, prox_studentcare, prox_eldercare, prox_cbd,
         prox_CHAS, prox_exercisepark, prox_gym, prox_hawker,
         prox_hospital, prox_library, prox_market, prox_parks,
         prox_pharmacy, prox_mall, prox_supermarket, 
         bus_within350m, preschool_within350m, 
         kindergarten_within350m, childcare_within350m, 
         prischool_within1km, studentcare_within350m,
         mrt_within350m, nrp_completed)
```

## Creating train/test data sets

We will train our predictive model using HDB resale transaction data from July 2024 to June 2025, to predict resale prices from July to September 2025. The following code chunk splits the dataset into training and test data sets.

```{r}
HDB_train <- HDBresale_sf %>%
  filter(month<='2025-06')
HDB_test <- HDBresale_sf %>%
  filter(month>'2025-06')
```

To prototype the models, we will use a smaller dataset to avoid excessively long run times for training and testing.

```{r}
set.seed(1234)
train_sample <- HDB_train %>%
  sample_n(5000)
```

## checking of overlapping points

when using GWmodel to calibrate explanatory or predictive models, it is important to ensure no overlapping point features

```{r}
overlapping_points<- train_sample %>%
  mutate(overlap=lengths(st_equals(., .))>1)
num_overlapping<- sum(overlapping_points$overlap)
num_overlapping
```

## spatial jittter

jthe code chunk below uses st_jitter of **sf** package to move point features by 1m to avoid overlapping point features

```{r}
train_sample<-train_sample %>%
  st_jitter(amount=1)
HDB_train<- HDB_train %>%
  st_jitter(amount = 1)
```

## Data sampling

`initial_sample` of **rsample** package allows splitting of data into test and training sets.

```{r}
set.seed(1234)
resale_split <- initial_split(train_sample, prop = 6.67/10)
train_data <- training(resale_split)
test_data <- testing(resale_split)
```

# Computing Correlation matrix

```{r}
#| fig-width: 10
HDB_train_nogeo <- HDB_train %>%
  st_drop_geometry() %>%
  select(c(-(month)))
corrplot::corrplot(cor(HDB_train_nogeo[, 2:17]), 
                   diag = FALSE, 
                   order = "AOE",
                   tl.pos = "td", 
                   tl.cex = 0.5, 
                   method = "number", 
                   type = "upper")
```

# Non Spatial Multiple Linear Regression 

```{r}
price_mlr <- lm(resale_price ~ floor_area_sqm + storey_order + 
         remaining_lease_months + prox_mrt + prox_cycling +
         prox_taxi + prox_studentcare + prox_eldercare + prox_cbd +
         prox_CHAS + prox_exercisepark + prox_gym + prox_hawker +
         prox_hospital + prox_library + prox_market + prox_parks +
         prox_pharmacy + prox_mall + prox_supermarket + 
         bus_within350m + preschool_within350m + 
         kindergarten_within350m + childcare_within350m + 
         prischool_within1km + studentcare_within350m +
         mrt_within350m + nrp_completed,
                data=HDB_train)
summary(price_mlr)
```

# References

1.  https://www.kaggle.com/datasets/muhdirshath/hospitals-in-singapore/data
2.  https://www.kaggle.com/datasets/sunnysharma432/singapore-malls-pois/data
3.  <https://www-jstor-org.libproxy.smu.edu.sg/stable/pdf/45389328>
4.  <https://personal.ntu.edu.sg/qfeng/BianChenFengLi20200804paper.pdf>
5.  <https://blogs.ubc.ca/zhuanlim/files/2022/07/Lim_Zhuan_A0224991X.pdf>
