---
title: "In Class Exercise 05"
format: html
author: "Erica Chan"
date: "27 September , 2025"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# Importing R packages

```{r}
pacman::p_load(tidyverse, sfdep, sf, tmap, spdep)
```

# Importing data into R

```{r}
hunan <- st_read(dsn = "data/geospatial", 
                 layer = "Hunan") %>% 
  st_transform(crs = 32650)
```

```{r}
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")
```
```{r}
hunan_GDPPC <- left_join(hunan, hunan2012) %>%
  select(1:4, 7, 15)
```

::: panel-tabset
# sfdep
Compute weights matrix with sfdep
```{r}
wm_q <- hunan_GDPPC %>%
  mutate(nb=st_contiguity(geometry),
         wt=st_weights(nb, 
                       style="W"),
         .before=1 #place any new columns created at the very beginning of the data frame.
  )
```

# spdep

Compute continuity matrix with spdep 
```{r}
wm_q_spdep <- poly2nb(hunan, queen=TRUE)
```
compute row standardised weight matrix with spdep
```{r}
rswm_q <- nb2listw(wm_q_spdep, 
                   style="W", 
                   zero.policy = TRUE)
rswm_q
```
:::
sfdep is an improvement on spdep because it can compute the weights matrix without intermediate steps, and it is easier to see the results in relation to the original table/locations without additional steps

# Global Moran's I

:::panel-tabset
# sfdep
Calculate Global Moran's I
```{r}
set.seed(1234) #ensure reproducibility of results
global_moran_perm(
  wm_q$GDPPC, 
  wm_q$nb, 
  wm_q$wt, 
  nsim=99 #100 simulations
  )
```
As p value <0.05, we reject the null hypothesis that the spatial distribution of GDP per capita resembles a random distirbution (ie are spatially independent). As the Moran's I statistic is greater than 0, we can infer that the GDP per capita is spatially clustered at the county level.

#spdep
```{r}
bperm= moran.mc(hunan_GDPPC$GDPPC, 
                listw=rswm_q, 
                nsim=999, 
                zero.policy = TRUE, 
                na.action=na.omit)
bperm
```
:::
sfdep allows the test to be performed using different columns within a single table

# Local Indicators of Spatial Association (LISA)

```{r}
set.seed(1234)
lisa <- wm_q %>%
  mutate(local_moran=local_moran(
    GDPPC, nb, wt, nsim=99),
    .before=1) %>%
  unnest(local_moran) 
```
```{r}
glimpse(lisa)
```
plot LISA 
```{r}
lisa <- lisa %>%
  mutate(
    LISA_cluster = ifelse(
      p_ii_sim < 0.05,
      as.character(mean), 
      "Insignificant"), 
    LISA_cluster = factor(
      LISA_cluster,
      levels = c("Insignificant", "Low-Low", "Low-High", "High-Low", "High-High")
    )
  )
```
```{r}
#| fig-width: 10
lisa_map <- tm_shape(lisa) + 
  tm_polygons(
    fill = "LISA_cluster",
    fill.scale = tm_scale_categorical(
      values = c(
        "grey80",      # Insignificant
        "blue",        # Low-Low
        "lightblue",   # Low-High
        "pink",        # High-Low
        "red"          # High-High
      )
    ),
    fill.legend = tm_legend(
      title = "LISA Cluster",
      position = tm_pos_in("left", "bottom"))
  ) +
  tm_borders() +
  tm_title("Local Moran's I Clusters (p < 0.05)")
lisa_map
```
This map shows statistically significant hot/cold spots.  

