---
title: "In-class_Ex04"
format: html
---

```{r}
pacman::p_load(sf, ggstatsplot, tmap, tidyverse, knitr, GWmodel)
```

read in the data and join the hunan and hunan2012 datasets

```{r}

hunan <- st_read(dsn = "data/geospatial", layer = "Hunan")
st_crs(hunan)
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")
hunan_sf <- left_join(hunan, hunan2012) %>% 
  select(1:4, 7, 15, 16, 31, 32)
```

Next, convert the data to sp format as GWmodel is built around sp format for handling spatial data in R

```{r}
hunan_sp <- hunan_sf %>%
  as_Spatial()
```

```{r}
bw_CV <- bw.gwr(GDPPC ~ 1,
                data = hunan_sp,
                approach = "CV",
                adaptive = FALSE, 
                kernel = "bisquare",
                longlat = T)
bw_CV
```

```{r}
bw_AIC <- bw.gwr(GDPPC ~ 1,
             data = hunan_sp, 
             approach = "AIC",
             adaptive = FALSE, 
             kernel = "bisquare",
             longlat = T
             )
bw_AIC
```

```{r}
gwstat <- gwss(data = hunan_sp,
               vars = "GDPPC", 
               bw = bw_AIC, 
               kernel = "bisquare",
               adaptive = FALSE, 
               longlat = T)
```

converting gwstat SDF data as a dataframe

```{r}
gwstat_df <- as.data.frame(gwstat$SDF)
```

next cbind used to append the dataframe onto the hunan_sf data frame

-   why? join uses a unique identifier to match rows

-   but gwstat does not have any identifier field like COUNTY.

    -   however, since gwstat was derived from the same dataframe and is sorted identically to hunan_sf (i.e., row sequence is the same), as long as no sorting has occurred, we can use cbind to join the two tables without a unique identifier

```{r}
hunan_gstat <- cbind(hunan_sf, gwstat_df)
```

```{r}

hunan_lm <- qtm(hunan_gstat, "GDPPC_LM") +
  tm_layout(
    legend.position = c("left", "bottom"),
    scale = 0.5) +
  tm_title("Distribution of geographically weighted mean")

hunan_lm
```
